Report for the Deep Learning Course Assignment 3

### 2016-12-16 practical_3/report/report_methenitis.tex
update report

@@

classif = OneVsRestClassifier(SVC(kernel='linear'))
classif.fit(features[train_set], batch_y[train_set]
classif.score(features[test_set], batch_y[test_set]
\end{verbatim}
Here, the reported accuracy\footnote{The accuracy is reported to half of the test set ($5000$ random samples without replacement due to slow training of the OneVsRestClassifier		)} of the one-vs-classifiers on the test set:
\begin{itemize}
\item Accuracy with flatten features: $0.59$.
\item Accuracy with fc1 features: $0.59$.
\item Accuracy with fc2 features: $0.703$.
\end{itemize}
It is natural that the accuracy is higher for later layers.
As it was earlier $fc2$ layer can distinguish better features of different classes.

@@

I used the following two loss functions:
\begin{enumerate}
\item $\mathcal{L} =  Y * \frac{1}{2} * d^2 + (1-Y) * max(margin - d^2, 0)$
\item $\mathcal{L} =  Y * d^2 + (1-Y) * max(margin - d^2, 0)$
\end{enumerate}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=10.2cm]{siamese.png}
\caption{\textbf{Experiment 1} Loss in the training (purple line) and the test set (yellow line) for $50000$ training steps, $margin = 1.0$, using the default \texttt{next\_batch} method and the loss function (1).}
\label{fig:siamese}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=10.2cm]{siamese-1.png}
\caption{\textbf{Experiment 2} Loss in the training (blue line) and the test set (light purple line) for $50000$ training steps, $margin = 1.0$, using the default \texttt{next\_batch} method and the loss function (2).
Loss in the training (purple line) and the test set (yellow line) for $50000$ training steps, $margin = 1.0$, using the default \texttt{next\_batch} method and the loss function (1).}
\label{fig:siamese-1}
\end{figure}

\begin{figure}[h!]
\centering
\includegraphics[width=10.2cm]{siamese-2.png}
\caption{\textbf{Experiment 3} Loss in the training (blue line) and the test set (light purple line) for $50000$ training steps, $margin = 1.0$, using the default \texttt{next\_batch} method and the loss function (2).
Loss in the training (yellow line) and the test set (light purple line) for $50000$ training steps, $margin = 1.0$, using the default \texttt{next\_batch} method and the loss function (2).}
\label{fig:siamese-2}
\end{figure}

@@

\begin{figure}[h!]
\centering
\includegraphics[width=6.2cm]{visualization-siamese-plain.pdf}
\includegraphics[width=6.2cm]{visualization-siamese-1.pdf}
\includegraphics[width=7.2cm]{visualization-siamese-3.pdf}
\caption{}
\label{fig:features}
\end{figure}

\subsection{One-vs-Rest Classifier}
Here, I present the accuracy of the one-vs-rest classifier on the features learned from in the final layer of the siamese network for all experiments
\begin{itemize}
\item \textbf{Experiment 1}: $0.476$
\item \textbf{Experiment 2}: $0.473$
\item \textbf{Experiment 3}: $\mathbf{0.682}$
\end{itemize}

\section{Transfer Learning}

@@

@@

%
%Using the same TSNE algorithm we present the $2$-dimensional representation of the feat
%\section{Conclusion}
%
%In this assignment I implemented a CNN and reported its performance over the CIFAR10 dataset using different parameters.
%I also implemented the siamese architecture which uses two versions of the same CNN to perform learning in order to distinguish between classes.
%The amount of training steps requires for the siamese are higher than the single CNN architecture.
%The time required for the experiments did not let me try more parameter values, therefore I cannot draw any important conclusions regarding the right parameters that should be used for the two different approaches.

>>>>

\subsection{Feature Visualization}
\label{sec:vis}

@@

classif = OneVsRestClassifier(SVC(kernel='linear'))
classif.fit(features[train_set], batch_y[train_set]
classif.score(features[test_set], batch_y[test_set]
\end{verbatim}
Here, the reported accuracy\footnote{The accuracy is reported to half of the test set ($5000$ random samples without replacement due to slow training of the OneVsRestClassifier		)} of the one-vs-classifiers on the test set:
\begin{itemize}
\item Accuracy with flatten features: $0.6305$.
\item Accuracy with fc1 features: $0.59$.
\item Accuracy with fc2 features: $\mathbf{0.7685}$.
\end{itemize}
It is natural that the accuracy is higher for later layers.
As it was earlier $fc2$ layer can distinguish better features of different classes.

@@

I used the following two loss functions:
\begin{enumerate}
\item $\mathcal{L} =  Y * \frac{1}{2} * d^2 + (1-Y) * max(margin - d^2, 0)$
\item $\mathcal{L} =  Y * d^2 + (1-Y) * max(margin - d^2, 0)$
\end{enumerate}
In the first the distance in similar label images is multiplied by half, while in the second is not.

@@

I construct 3 experiments using both the \texttt{next\_batch} ways (default \& alternative) and the two loss functions.
\begin{description}
\item[Experiment 1]{Using the loss function (1) and the default creation of batches}
\item[Experiment 2]{Using the loss function (2) and the default creation of batches}
\item[Experiment 3]{Using the loss function (2) and the alternative creation of batches}
\end{description}
Figure~\ref{fig:siamese} presents the loss in the training and the test set for $50000$ training steps for the setting of the experiment 1.
Loss in the test set is computed every $1000$ steps.
We see that the loss in the training set in decreasing as the training steps increase.
The same occurs for the loss in the test set.
\begin{figure}[h!]
\centering
\includegraphics[width=10.2cm]{siamese.png}
\caption{\textbf{Experiment 1} Loss in the training (purple line) and the test set (yellow line) for $50000$ training steps, $margin = 1.0$, using the default \texttt{next\_batch} method and the loss function (1).}
\label{fig:siamese}
\end{figure}
Figure~\ref{fig:siamese-1} presents the loss in the training and the test set for $50000$ training steps for the setting of the experiment 2.
Here the scaling of the loss is different as a result of the different loss function used in experiment 2.
\begin{figure}[h!]
\centering
\includegraphics[width=10.2cm]{siamese-1.png}
\caption{\textbf{Experiment 2} (Different Loss Functions) Loss in the training (blue line) and the test set (light purple line) for $50000$ training steps, $margin = 1.0$, using the default \texttt{next\_batch} method and the loss function (2).
Loss in the training (purple line) and the test set (yellow line) for $50000$ training steps, $margin = 1.0$, using the default \texttt{next\_batch} method and the loss function (1).}
\label{fig:siamese-1}
\end{figure}
Figure~\ref{fig:siamese-2} presents the loss in the training and the test set for $50000$ training steps for the setting of the experiment 3.
Here I compare between two different ways of batch sampling between experiment 2 and 3.
The same loss function in used both experiments.
It is easy to see that using the alternative batch sampling, the loss is decreasing faster and results in lower loss in the test set as well.
In the next section I show that the features retained with the alternative batch sampling increase the accuracy of the one-vs-rest classifier.
\begin{figure}[h!]
\centering
\includegraphics[width=10.2cm]{siamese-2.png}
\caption{\textbf{Experiment 3} (Default vs Alternative Batch Creation) Loss in the training (blue line) and the test set (light purple line) for $50000$ training steps, $margin = 1.0$, using the default \texttt{next\_batch} method and the loss function (2).
Loss in the training (yellow line) and the test set (light purple line) for $50000$ training steps, $margin = 1.0$, using the default \texttt{next\_batch} method and the loss function (2).}
\label{fig:siamese-2}
\end{figure}

@@

In the previous section I set up three experiments and presented the loss during training in the training set.
In this section I illustrate the features in the 2-dimensional space as these transformed by TSNE (as this was implemented in Section~\ref{sec:vis}.).

\begin{figure}[h!]
\centering
\includegraphics[width=6.2cm]{visualization-siamese-plain.pdf}
\includegraphics[width=6.2cm]{visualization-siamese-1.pdf}
\includegraphics[width=7.2cm]{visualization-siamese-3.pdf}
\caption{Visualization of the $2D$ features generated out of the whole test set. Experiment 1 (top left), experiment 2 (top right), experiment 3 (bottom).}
\label{fig:features}
\end{figure}

\subsection{One-vs-Rest Classifier}

Here, I present the accuracy of the one-vs-rest classifier on the features learned from in the final layer of the siamese network for all experiments.

\begin{itemize}
\item \textbf{Experiment 1}: $0.476$
\item \textbf{Experiment 2}: $0.473$
\item \textbf{Experiment 3}: $\mathbf{0.682}$
\end{itemize}

Using the different loss function with the default batch sampling method did not help the accuracy obtained.
However, in experiment 3 I used the alternative batch sampling function which was the main reason for the large improvement in the accuracy.

\clearpage

\section{Task 3: Transfer Learning}

@@

In this section I present the results of my implementation for the feature extraction of the third task.
I used the last 3 layers of the convNet inference method from the first task of the assignment.
I also used stopGradient method of tensorflow to disconnect vgg pretrained network from convNet and the training updates.

@@

\subsection{Performance of transfer learning on Cifar10}
\begin{description}
\item[Feature extraction:]{$0.6743$}
\item[Retraining:]{$0.8505$}
\item[Refining-$100$:]{$0.8441$}
\item[Refining-$1000$:]{$0.8461$}
\item[Refining-$2500$:]{$\mathbf{0.8510}$}
\end{description}

\section{Conclusion}

In this assignment I implemented a CNN and reported its performance over the CIFAR10 dataset using different parameters.
I also implemented the siamese architecture which uses two versions of the same CNN to perform learning in order to distinguish between classes.
I showed that using the alternative batch sampling procedure the loss is reduced and the features retained result in better accuracy when an one-vs-rest classifier is used.
The amount of training steps requires for the siamese are higher than the single CNN architecture.
Last, I implemented the third task of the assignment where we import pre-trained layers from VGG-16 and apply learning on Cifar10 dataset.
